Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-08 arxiv/proof_examples.tex
restructure B.2

%\subsubsection{Proof for Sample Imbalance}
We provide the proof of Proposition \ref{cor_MTL_loss}. Similar to the proof of Theorem \ref{thm_many_tasks}, our proof is divided into the following parts:
\begin{itemize}
\item First, we reduce the optimization objective to another function $g(x)$ that only depends on the sample covariance matrices.
\item Second, we show that the global minimizer of $g(x)$ is approximately equal to $1$.
\item Finally, we use the above result and Theorem \ref{thm_main_RMT} to prove the proposition.
\end{itemize}

\iffalse
%Our key insight is a bias-variance decomposition of the expected prediction loss of $\hat{\beta}_2^{\MTL} = \hat{B} W_2$.
%Using the local optimality condition of $B$, we obtain that
%\begin{align*}
%	 \hat{B}(W_1, W_2) &= (W_1^2 X_1^{\top}X_1 + W_2^2 X_2^{\top}X_2)^{-1} (W_1 X_1^{\top}Y_1 + W_2 X_2^{\top}Y_2)\\
%	&= \frac{1}{W_2} \left( \frac{W_1^2}{W_2^2}  X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(\frac{W_1}{W_2} X_1^{\top}Y_1 + X_2^{\top}Y_2\right) \\
%	&= \frac{1}{W_2}\left[\beta_2 + \left(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}\bigbrace{X_1^{\top}X_1\left(\frac{W_1}{W_2}\beta_1 - \frac{W_1^2}{W_2^2} \beta_2\right) + \left(\frac{W_1}{W_2} X_1^{\top}\varepsilon_1 + X_2^{\top}\varepsilon_2\right)}\right].
%\end{align*}
Recall that $\hat{\beta}_i^{\MTL} = \hat{B} W_i$. Therefore, we have
\begin{align}
	\exarg{\epsilon_1, \epsilon_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2}
	=&~ \frac{W_1^2}{W_2^2} \bignorm{\Sigma_2^{1/2}(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_1 - \frac{W_1}{W_2} \beta_2)}^2 \label{eq_bias_2task} \\
			&+~  \sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }. \label{eq_var_2task}
\end{align}
equation \eqref{eq_bias_2task} is the bias of $\hat{\beta}_t^{\MTL}$ and
equation \eqref{eq_var_2task} is the variance of $\hat{\beta}_t^{\MTL}$.
%minus the variance of $\hat{\beta}_t^{\STL}$, which is always negative.
Comparing the above to single-task learning, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma_2 (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

% \todo{derive the test loss} Expectation was given in equation \eqref{eq_bias_2task} and equation \eqref{eq_var_2task}.
 \fi

\paragraph{Part $1$: the reduction.}
Under the two-task setting of Corollary \ref{cor_MTL_loss}, the optimization objective \eqref{eq_mtl} becomes

@@

		f(A_1,A_2, B) =   \bignorm{X^{(1)} B A_1 - Y^{(1)}}^2+ \bignorm{X^{(2)} B A_2 - Y^{(2)}}^2, \label{eq_mtl_2task_cov}

@@

	where $B \in \real^{p}$ and $A_1, A_2\in \R$ because $r=1$ in the two-task case. % for $1 < r < t$ by our assumption.
	Without loss of generality, we assume that both $A_1$ and $A_2$ are nonzero, since otherwise we can add a small perturbation $\delta$ to them and later remove this perturbation by taking $\delta\to 0$ in the final result. Again using the local optimality condition ${\partial f}/{\partial B} = 0$, for any $A_1$ and $A_2$ we obtain the optimal $\hat{B}$ as follows

@@

 Now inserting $\hat B$ back into equation \eqref{eq_mtl_2task_cov}, we obtain an objective that only depends on $x:=A_1/A_2$ as follows %\HZ{$A$ has been used to denote the output layers. Could you replace $A$ with another symbol (say $x$)?}

@@

		 &\left. + x X^{(2)}\hat \Sigma(x)^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2, 

@@

	where we denote $ \hat \Sigma(x):= x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$ We can calculate the partial expectation of $g(x)$ over the randomness in $\epsilon_1$ and $\epsilon_2$:

@@

&\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta_1,\beta_2} \nonumber\\

@@

&=(\beta_1-x\beta_2)^\top (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1} (X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)}  (\beta_1-x\beta_2)\\

@@

&=(\beta_1-x\beta_2)^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-x\beta_2).

@@

		&= (\beta_1-x\beta_2)^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-x\beta_2)  \\

@@

		& = (\beta_1-x\beta_2)^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-x\beta_2) +\sigma^2(n_1+n_2-p).

@@

For the random-effects model, we can further take the partial expectation over $\beta_1$ and $\beta_2$: 

@@

%and in the second step we used  %Using $(X^{(2)})^\top X^{(2)}=\hat\Sigma- A^2 (X^{(1)})^\top X^{(1)}$, we can further simplify that
% & (X^{(2)})^\top X^{(2)} \hat\Sigma(x)^{-1}(X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} + A^2   (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1}(X^{(2)})^\top X^{(2)} \hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)}  \\
% &=  (X^{(2)})^\top X^{(2)} \hat\Sigma(x)^{-1}(X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} + A^2  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1}(X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1}(X^{(2)})^\top X^{(2)}  \\
% & =   (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1}(X^{(2)})^\top X^{(2)} .

Using the concentration estimates in Lemma \ref{largedeviation}, we can control the difference between $g(x)$ and its partial expectation. %$\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta_1,\beta_2}$.
 \begin{claim}\label{claim_largedev1}
 In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
$$\left| g(x) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(x) \mid X_1, X_2 }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right) $$
for any small constant $c>0$.
	% \HZ{use consistent $e_1, e_2, e$, etc. Why not just use $\e$? Don't create too many different symbols. It's bad style..}
 \end{claim}
 \begin{proof}
By Fact \ref{lem_minv} (iii), we have that with high probability,
 \be\label{op_X12}
\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, 
 \ee
 and
 \be\label{op_Sig1}
\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
 \ee
where we used that $1+\tau\le \rho_1=n_1/p\le \tau^{-1}$ and $1+\tau\le \rho_2=n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

\paragraph{Part $2$: approximating the global minimum.}
	Let $\hat{x}$ denote the minimizer of $g(x)$. 
	We show that in the setting of Corollary \ref{cor_MTL_loss}, $\hat x$ is close to 1.% This will be used to calculate the predication loss $\te(\hat{\beta}_2^{\MTL})$ loss.
%Our main observation is that the optimum of $\val(v)$ is approximately $1$.This allows us to simplify the bias and variance bounds significantly.
%Furthermore, under certain choice of parameters, we can simplify  
%More precisely, we have the following claim.
%Later on, we will combine it with Lemma \ref{prop_model_shift_tight} together to analyze the bias-variance tradeoff.

\begin{claim}\label{lem_hat_v}
%Suppose the assumptions of Lemma \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order.
 In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
%There exists a constant $C>0$ such that
	\be\label{hatw_add1}|\hat x -1|\le  \frac{2d^2}{\kappa^2} + p^{-1/4+c}. 
	\ee
	for any small constant $c>0$.
	%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{claim}

@@

\paragraph{Part $3$: Applying Theorem \ref{thm_main_RMT}.}
We provide the predication loss using the optimal $\hat B$ in equation \eqref{eq_Bhat_2task} and optimal $\hat x=\hat x_1/\hat x_2$:
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left\|\Sigma_2^{1/2} \left( \hat B \hat x_2 - \beta_2\right)\right\| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta_1-\hat x^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
%where we denoted $$\hat \Sigma:= \hat x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
To illustrate the idea of bias-variance decomposition, we assume that $\hat x$ is independent of $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$ for now. Then we can calculate the partial expectation of $L(\hat{\beta}_2^{\MTL})$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
 \begin{align}
	\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2,\beta_1,\beta_2}  &=  \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 \nonumber\\ %\label{eq_bias_2task} 
			& + \sigma^2\cdot \bigtr{ X^{(2)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(2)})^\top} \nonumber\\
			&+ \sigma^2\cdot \hat x^2 \bigtr{ X^{(1)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(1)})^\top}\nonumber\\
			%+\sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } \nonumber\\ 
			&= \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 +\sigma^2  \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).\nonumber
			%\label{eq_var_2task}
\end{align}
This gives a bias-variance decomposition of the MTL prediction loss: the first term on the right-hand side is the bias term, while the second term is the variance term. For the random-effects model, we further take the partial expectation over $\beta_1$ and $\beta_2$: 
 \begin{align}
	& \exarg{\epsilon^{(1)}, \epsilon^{(2)},\beta_1,\beta_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2} \nonumber\\
	&	=  \hat x^2 \left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} }\nonumber \\
	&+\sigma^2 \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
			\label{eq_var_2task}
\end{align}
One can compare the above expression with the predication loss for STL, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma^{(2)} (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over the second term on the right-hand side of equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

@@

% \todo{switch the $a_3$ and $a_4$}
Finally, we are ready to complete the proof of Proposition \ref{cor_MTL_loss}.
We derive the variance term $\sigma^2  \tr(\hat\Sigma^{-1})$ and the bias term $d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right]$ using Theorem \ref{thm_main_RMT}.
%For the isotropic model in Section \ref{sec_similarity}, we actually have an easier and sharper bound than Lemma \ref{thm_model_shift}. %as follows.
%\begin{lemma}\label{prop_model_shift_tight}
%		In the setting of Lemma \ref{thm_model_shift}, assume that $\Sigma_1 =\id$,
%		$\beta_2$ is a random vector that has i.i.d. entries with mean $0$, variance $\kappa^2$ and finite moments up to any order, and $(\beta_1-\beta_2)$ is a random vector that is independent of $\beta_2$ and has i.i.d. entries with mean $0$, variance $d^2$ and finite moments up to any order. Denote
%		$\Delta^\star_{\bias} := \bigbrace{(1 - \hat{v})^2 \kappa^2 + d^2} \bigtr{\Pi(\hat v)}$, where $\Pi$ was defined in equation \eqref{defnpihat}.	Then we have with high probability,
%		\begin{align*}
%			\te(\hat{\beta}_t^{\MTL}) \le \te(\hat{\beta}_t^{\STL})\quad \text{ when: }\ \  & \Delta_{\vari} \ge  (\al_+^2(\rho_1)+\oo(1))\cdot  \Delta^\star_{\bias} \ , \\
%			\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})\quad \text{ when: }\ \  & \Delta_{\vari} \le (\al_-^2(\rho_1)-\oo(1))\cdot  \Delta^\star_{\bias}\ ,
%		\end{align*}
%		where $\Delta_{\vari}$ was defined in equation \eqref{Deltavarv}.
%\end{lemma}

>>>>

We provide the proof of Corollary \ref{cor_MTL_loss}. The proof is similar to that of Theorem \ref{thm_many_tasks}.
First, we reduce the optimization objective to another function $g(x)$ that only depends on the sample covariance matrices.
Second, we show that the global minimizer of $g(x)$ is approximately equal to $1$.
Finally, we use the above result and Theorem \ref{thm_main_RMT} to prove the proposition.

In the setting of Corollary \ref{cor_MTL_loss}, the optimization objective \eqref{eq_mtl} is equal to

@@

		f(A, B) =   \bignorm{X^{(1)} B A_1 - Y^{(1)}}^2+ \bignorm{X^{(2)} B A_2 - Y^{(2)}}^2, \label{eq_mtl_2task_cov}

@@

	where $B \in \real^{p}$ and $A = [A_1, A_2]\in \real^{2}$ because the width of $B$ is one.
We assume that $A_1$ and $A_2$ are both nonzero.
Otherwise, we add a tiny amount of perturbation $\delta$ to them and the result remains the same.
Using the local optimality condition $\frac{\partial f}{\partial B} = 0$, we have that $\hat{B}$ satisfies the following

@@

Let denote $\hat \Sigma(x):= x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)}$.
Applying $\hat B$ to equation \eqref{eq_mtl_2task_cov}, we obtain an objective that only depends on $x:=A_1/A_2$ as follows %\HZ{$A$ has been used to denote the output layers. Could you replace $A$ with another symbol (say $x$)?}

@@

		 &\left. + x X^{(2)}\hat \Sigma(x)^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2.

@@

The conditional expectation of $g(x)$ over $\epsilon_1$ and $\epsilon_2$ is

@@

\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta_1,\beta_2}

@@

=&(\beta_1-x\beta_2)^\top (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1} (X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)}  (\beta_1-x\beta_2)\\

@@

=&(\beta_1-x\beta_2)^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-x\beta_2).

@@

		=& (\beta_1-x\beta_2)^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-x\beta_2)  \\

@@

		=& (\beta_1-x\beta_2)^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-x\beta_2) +\sigma^2(n_1+n_2-p).

@@

For the random-effects model, we take the expectation over $\beta_1$ and $\beta_2$:

@@

\paragraph{Part $1$: approximating the global minimum.}
	Let $\hat{x}$ denote the minimizer of $g(x)$.
	We show that in the setting of Corollary \ref{cor_MTL_loss}, $\hat x$ is close to 1.% This will be used to calculate the predication loss $\te(\hat{\beta}_2^{\MTL})$ loss.
%Our main observation is that the optimum of $\val(v)$ is approximately $1$.This allows us to simplify the bias and variance bounds significantly.
%Furthermore, under certain choice of parameters, we can simplify
%More precisely, we have the following claim.
%Later on, we will combine it with Lemma \ref{prop_model_shift_tight} together to analyze the bias-variance tradeoff.

\begin{claim}\label{lem_hat_v}
%Suppose the assumptions of Lemma \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order.
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability,
%There exists a constant $C>0$ such that
	\be\label{hatw_add1}|\hat x -1|\le  \frac{2d^2}{\kappa^2} + p^{-1/4+c}.
	\ee
%	for any small constant $c>0$.
	%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{claim}

@@

\paragraph{Part $2$: reducing the random-effects model to Theorem \ref{thm_main_RMT}.}
Using the concentration estimates in Lemma \ref{largedeviation}, we first show that $g(x)$ and its conditional expectation over $X^{(1)}, X^{(2)}$ are close. %$\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta_1,\beta_2}$.
 \begin{claim}\label{claim_largedev1}
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability
$$\left| g(x) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(x) \mid X_1, X_2 }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right)$$.
%for any small constant $c>0$.
	% \HZ{use consistent $e_1, e_2, e$, etc. Why not just use $\e$? Don't create too many different symbols. It's bad style..}
 \end{claim}
 \begin{proof}
By Fact \ref{lem_minv} (iii), we have that with high probability,
 \be\label{op_X12}
\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, 
 \ee
 and
 \be\label{op_Sig1}
\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
 \ee
where we used that $1+\tau\le \rho_1=n_1/p\le \tau^{-1}$ and $1+\tau\le \rho_2=n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

We provide the predication loss using the optimal $\hat B$ in equation \eqref{eq_Bhat_2task} and optimal $\hat x=\hat x_1/\hat x_2$:
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left\|\Sigma_2^{1/2} \left( \hat B \hat x_2 - \beta_2\right)\right\| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta_1-\hat x^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
%where we denoted $$\hat \Sigma:= \hat x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
To illustrate the idea of bias-variance decomposition, we assume that $\hat x$ is independent of $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$ for now. Then we can calculate the partial expectation of $L(\hat{\beta}_2^{\MTL})$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
 \begin{align}
	\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2,\beta_1,\beta_2}  &=  \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 \nonumber\\ %\label{eq_bias_2task} 
			& + \sigma^2\cdot \bigtr{ X^{(2)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(2)})^\top} \nonumber\\
			&+ \sigma^2\cdot \hat x^2 \bigtr{ X^{(1)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(1)})^\top}\nonumber\\
			%+\sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } \nonumber\\ 
			&= \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 +\sigma^2  \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).\nonumber
			%\label{eq_var_2task}
\end{align}
This gives a bias-variance decomposition of the MTL prediction loss: the first term on the right-hand side is the bias term, while the second term is the variance term. For the random-effects model, we further take the partial expectation over $\beta_1$ and $\beta_2$: 
 \begin{align}
	& \exarg{\epsilon^{(1)}, \epsilon^{(2)},\beta_1,\beta_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2} \nonumber\\
	&	=  \hat x^2 \left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} }\nonumber \\
	&+\sigma^2 \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
			\label{eq_var_2task}
\end{align}
One can compare the above expression with the predication loss for STL, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma^{(2)} (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over the second term on the right-hand side of equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

@@

\paragraph{Part $3$: Applying Theorem \ref{thm_main_RMT}.}
Finally, we are ready to complete the proof of Corollary \ref{cor_MTL_loss}.
We derive the variance term $\sigma^2  \tr(\hat\Sigma^{-1})$ and the bias term $d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right]$ using Theorem \ref{thm_main_RMT}.
%For the isotropic model in Section \ref{sec_similarity}, we actually have an easier and sharper bound than Lemma \ref{thm_model_shift}. %as follows.
%\begin{lemma}\label{prop_model_shift_tight}
%		In the setting of Lemma \ref{thm_model_shift}, assume that $\Sigma_1 =\id$,
%		$\beta_2$ is a random vector that has i.i.d. entries with mean $0$, variance $\kappa^2$ and finite moments up to any order, and $(\beta_1-\beta_2)$ is a random vector that is independent of $\beta_2$ and has i.i.d. entries with mean $0$, variance $d^2$ and finite moments up to any order. Denote
%		$\Delta^\star_{\bias} := \bigbrace{(1 - \hat{v})^2 \kappa^2 + d^2} \bigtr{\Pi(\hat v)}$, where $\Pi$ was defined in equation \eqref{defnpihat}.	Then we have with high probability,
%		\begin{align*}
%			\te(\hat{\beta}_t^{\MTL}) \le \te(\hat{\beta}_t^{\STL})\quad \text{ when: }\ \  & \Delta_{\vari} \ge  (\al_+^2(\rho_1)+\oo(1))\cdot  \Delta^\star_{\bias} \ , \\
%			\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})\quad \text{ when: }\ \  & \Delta_{\vari} \le (\al_-^2(\rho_1)-\oo(1))\cdot  \Delta^\star_{\bias}\ ,
%		\end{align*}
%		where $\Delta_{\vari}$ was defined in equation \eqref{Deltavarv}.
%\end{lemma}

\begin{proof}[Proof of Corollary \ref{cor_MTL_loss}]
%The proof of Lemma \ref{prop_model_shift_tight} is similar to that for Lemma \ref{thm_model_shift}, except that we can replace equation \eqref{bounddelta-} with a tighter bound. We only describe the main difference in the following proof.
%For any fixed $v\in \R$, $\beta_1 - v\beta_2$ is a random vector with i.i.d. entries with mean $0$ and variance $(1-v)^2\kappa^2 + d^2$. Then using the concentration result, Lemma \ref{largedeviation}, we get that for any constant $\e>0$, 
%\begin{align}
%&\left|\delta_{\bias}(v)- [(1-v)^2\kappa^2 + d^2]\tr (\cal K^\top \cal K)\right| \nonumber\\
%&=  \left|(\beta_1 - {v} \beta_2)^\top \cal K^\top \cal K (\beta_1 - {v} \beta_2) - [(1-v)^2\kappa^2 + d^2]\tr (\cal K^\top \cal K)\right| \nonumber\\
%&\le p^\e [(1-v)^2\kappa^2 + d^2] \left\{\tr\left[(\cal K^\top \cal K)^2\right]\right\}^{1/2} \lesssim p^{1/2+\e} [(1-v)^2\kappa^2 + d^2] \quad \text{w.h.p.},\label{anotherbeta}
%\end{align}
%where $\delta_{\bias}(v)$ was defined in equation \eqref{revise_deltabias}, $\cal K$ was defined as $\cal K:=v\Sigma_2^{1/2}({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1$, and in the last step we used $\|\cal K\|=\OO(1)$ by Fact \ref{lem_minv}. Now for $\tr (\cal K^\top \cal K)$, we rewrite it as
%\begin{align*}
%v^2 \tr \left[ ({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}(X_1^\top X_1)^2 \right].
%\end{align*}
%Bounding $(X_1^\top X_1)^2=(Z_1^\top Z_1)^2$ 
Using \eqref{lem_cov_shift_eq}, we obtain that 
\be\label{eq_var111}\tr(\hat\Sigma^{-1}) = \tr\left[ \left((X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)}\right)^{-1}\right]=\bigtr{\frac{(a_1\id_p+a_2\id_p)^{-1}}{n_{1}+n_2} }+\OO(p^{-c_\varphi})\ee
with high probability. Solving equation \eqref{eq_a12extra} with $\lambda_i\equiv 1$, $1\le i\le p$, we get that  
	\begin{align}
		 a_1 = \frac{\rho_1(\rho_1 + \rho_2 - 1)}{(\rho_1 + \rho_2)^2} ,\quad
		& a_2 = \frac{\rho_2(\rho_1 + \rho_2 - 1)}{(\rho_1 + \rho_2)^2} . \label{simplesovlea12}
			\end{align}
Inserting it into \eqref{eq_var111}, we obtain that 
\be\label{eq_var112}\tr(\hat\Sigma^{-1})  = \frac{p}{n_1+n_2} \cdot \frac{\rho_1+\rho_2}{\rho_1+\rho_2-1}+\OO(p^{-c_\varphi})=  \frac{1}{\rho_1+\rho_2-1}+\OO(p^{-c_\varphi})\ee
with high probability.
